### 2010-04-01 papers/icfp-2010-desc/paper_desc_levitation.tex
Desc@ICFP: inline figures

In order to capture a notion of higher-order induction, we need to
extend our universe of descriptions. This consists in adding a code
$\DHindx{\!}{\!}$ that takes an indexing set $\V{H}$. The code and its
interpretation are presented in
Figure~\ref{fig:hindx_desc}. Intuitively, $\DHindx{\!}{\!}$ uses the
elements of $H$ to index as many recursive arguments.

\begin{figure}

@@

\caption{Higher-order universe of descriptions}
\label{fig:hindx_desc}

\end{figure}

@@

As shown in Figure~\ref{fig:desc-levitate}, we are now able to
describe our universe of datatypes.  The $\DUnit$ and $\DIndx{\!}$
cases remain unchanged, as expected. We successfully describe the
$\DSigma{\!}{}$ case, by a simple appeal to the higher-order induction
on $\V{S}$. The $\DHindx{\!}{\!}$ case consists in packing an $\V{H}$
in $\Set$ with a recursive argument.

@@

\begin{figure}

\caption{Levitating description of $\Desc$}
\label{fig:desc-levitate}

\end{figure}

@@

Our reader might be slightly confused to learn that the trick is
visible in the definition of $\DescD$
(Fig.~\ref{fig:desc-levitate}). Or rather, it is made invisible by
careful usage of type propagation. Indeed, let us try to elaborate
this term down to the low-level type theory. The $\EnumT{[ \ldots ]}$
construct elaborates to a finite set $\V{E}$ in $\EnumU$, inhabited by
the codes. Then, we can type-check the case definition, between square
brackets $[ \ldots ]$. This term $\V{f}$ is pushed into the type
$\EnumT{\V{E}} \To \Mu{\DescD}$. This corresponds to a finite function
definition, as presented in Figure~\ref{fig:type-checking}.

@@

\[
\Rule{\Gamma \Vdash \propag{\push{\V{t}}{\spiD{\V{e}}}}
                           {\V{t'}}}
     {\Gamma \Vdash \begin{array}{@{}l}
                        \propag{\push{\V{t}}{\EnumT{\V{e}} \To \Desc}}
                               {\\ \PLAM{\V{x}}{(\EnumT{\V{e}})} \switchD{\V{e}}{\V{t'}}{\V{x}}}
                    \end{array}
     }\;\mbox{$\V{t}$ is $[]$ or $[\_,\_]$}
\]

>>>>

In order to capture a notion of higher-order induction, we need to
extend our universe of descriptions. This consists in adding a code
$\DHindx{\!}{\!}$ that takes an indexing set $\V{H}$. Intuitively,
$\DHindx{\!}{\!}$ uses the elements of $H$ to index as many recursive
arguments:

@@

@@

We are now able to describe our universe of datatypes:

The $\DUnit$ and $\DIndx{\!}$ cases remain unchanged, as expected. We
successfully describe the $\DSigma{\!}{}$ case, by a simple appeal to
the higher-order induction on $\V{S}$. The $\DHindx{\!}{\!}$ case
consists in packing an $\V{H}$ in $\Set$ with a recursive argument.

@@

@@

Our reader might be slightly confused to learn that the trick is
visible in the definition of $\DescD$. Or rather, it is made invisible
by careful usage of type propagation. Indeed, let us try to elaborate
this term down to the low-level type theory. The $\EnumT{[ \ldots ]}$
construct elaborates to a finite set $\V{E}$ in $\EnumU$, inhabited by
the codes. Then, we can type-check the case definition, between square
brackets $[ \ldots ]$. This term $\V{f}$ is pushed into the type
$\EnumT{\V{E}} \To \Mu{\DescD}$. This corresponds to a finite function
definition, as presented in Figure~\ref{fig:type-checking}.

@@

\[
\Rule{\Gamma \Vdash \propag{\push{\V{t}}{\spiD{\V{e}}}}
                           {\V{t'}}}
     {\Gamma \Vdash \propag{\push{\V{t}}{\EnumT{\V{e}} \To \Desc}}
                           {\PLAM{\V{x}}{(\EnumT{\V{e}})} \switchD{\V{e}}{\V{t'}}{\V{x}}}
     }\;\mbox{$\V{t}$ is $[]$ or $[\_,\_]$}
\]
